---
title: 'Assignment 6: Sobol Sensitivity Analysis'
author: "Mia Guarnieri, Lauren Harris"
date: "2023-05-23"
output:
  html_document:
    toc: yes
    toc_float: yes
    theme: paper
    code_folding: hide
    smooth_scroll: yes
    collapsed: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(deSolve)
library(sensitivity)
library(here)
```

### Source the function and run it for 300 years

```{r sen}
source(here("R", "forest_growth.R"))

#initial forest size (carbon)
C <- 10

#initial parameters
K <- 250
r <- 0.01
g <- 2
closure <- 50

#run the model for 300 years using ODE solver

#set up parameters
simtimes <- seq(from = 1, to = 300)
parms <- list(r = r, K = K, g = g, closure = closure)


result1 <- ode(y = C, times = simtimes, func = forest_growth, parms = parms)

#turn into a df for plotting
res1_df <- as.data.frame(result1)
colnames(res1_df) = c("time", "forest_size")

#plot
ggplot(res1_df, aes(time, forest_size)) + 
  geom_point() +
  labs(x = "Time (Years)",
       y = "Forest Size (kg/C)")

```

### Run sobol sensitivity analysis

Exploring estimated maximum forest size variation with the parameters:
- pre canopy closure growth rate (r)
- post-canopy closure growth rate (g)
- canopy closure threshold (closure) and carrying capacity(K)

```{r echo=FALSE, results = "hide"}
#set up parameter sets
np <- 2000 #number of parameters

#set 1
r <- rnorm(mean = 0.01, sd = 0.001, n = np)
g <- rnorm(mean= 2, sd = 0.2, n = np)
closure <- rnorm(mean = 50, sd = 5, n = np)
K <- rnorm(mean = 250, sd = 25, n = np)
X1 <- cbind.data.frame(r = r, g = g, closure = closure, K = K)

#set 2
r <- rnorm(mean = 0.01, sd = 0.001, n = np)
g <- rnorm(mean= 2, sd = 0.2, n = np)
closure <- rnorm(mean = 50, sd = 5, n = np)
K <- rnorm(mean = 250, sd = 25, n = np)
X2 <- cbind.data.frame(r = r, g = g, closure = closure, K = K)

# turn any negative values into 0
X1 <- X1 %>% 
  map_df(pmax, 0.0)
X2 <- X2 %>% 
  map_df(pmax, 0.0)

#create sobol object
sens_C = sobolSalt(model = NULL, X1, X2, nboot = 300)

#add column names
colnames(sens_C$X) = c("r", "g", "closure", "K")

#get results for 300 years (evaluating every year)

#define a wrapper function to run solver, compute metrics, and send back results for each parameter
c_wrapper <- function(r, g, closure, K, Cinit, times, func) {
    parms <- list(r = r, g = g, closure = closure, K = K)
    result <- ode(y = Cinit, times = times, func = func, parms = parms) 
  
    #turn into dataframe and get max value
    result <- as.data.frame(result)
    colnames(result) = c("time", "C")
    max_C <- max(result$C)
    
    return(list(max_C = max_C))
}

#use pmap to apply wrapper to all parameter sets

times = 1:300
Cinit = 10

allresults <- as.data.frame(sens_C$X) %>% 
  pmap(c_wrapper, Cinit = Cinit, times = times, func = forest_growth)

```
```{r}
# extract out results from pmap into a data frame
allres <- allresults %>% 
  map_dfr(`[`,c("max_C"))

# create boxplot
ggplot(allres, aes(y = max_C)) +
  geom_boxplot(color = "forestgreen") +
  labs(y = "Maximum forest size after 300 years (kg/C)") +
  theme_minimal() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```


### Calculate Sobol indices

```{r}
sens_C_max = sensitivity::tell(sens_C, allres$max_C)

# first-order indices (main effect without co-variance)
rownames(sens_C_max$S) = c("r", "g", "closure", "K")
sens_C_max$S

# total sensitivity index 
rownames(sens_C_max$T) = c("r", "g", "closure", "K")
sens_C_max$T
```

## Discussion
We find that variability in r, or pre canopy closure growth rate, has the greatest affect on variability in maximum forest size after 300 years. Variability in g (post canopy closure growth rate) has the second greatest affect on variability in maximum forest size, followed by the canopy closure threshold ("closure") and then carrying capacity (K). This means that the most important factors affecting the size of a forest (according to this analysis) are the growth rates of the forest rather than carrying capacity or the canopy closure threshold. Growth rates of trees may be affected by resource availability, such as precipitation and soil nutrients, which can both be impacted by climate change as increasing temperature and weather extremes reduce the suitability of many habitats globally. Under the scenario analyzed, ensuring that the conditions remain as optimal as possible for tree growth will ultimately allow land managers to maximize forest size.
